---
title: "Perform permutation test for differences in gene expression"
output: html_document
---

Set a seed to make the permutation analysis reproducible.

```{r seed}
set.seed(9)
```

Import the mice weights.

```{r import-data}
weights <- read.delim("../data/weights.txt",
                      stringsAsFactors = FALSE)
head(weights)
```

Calculate the observed difference in weight between wild type and transgenic
mice.

```{r observed}
# Calculate the mean difference in weight between Wild type and Transgenic mice.
#
# mouse - character vector of type of mouse: Wild type, Transgenic
# grams - numeric vector of mouse weights (grams)
get_diff <- function(mouse, grams) {
  stopifnot(length(mouse) == length(grams))
  mean(grams[mouse == "Transgenic"]) / mean(grams[mouse == "Wild type"])
}

diff_observed <- get_diff(weights$mouse, weights$grams)
diff_observed
```

Perform 1000 permutations and calculate the mean difference in each permuted
data set.

```{r permute}
n_perms <- 1000
permutations <- replicate(n = n_perms, expr = sample(weights$mouse),
                          simplify = FALSE)
diff_permutations <- vapply(permutations, get_diff, grams = weights$grams,
                            numeric(1))
hist(diff_permutations)
abline(v = diff_observed, col = "red")
```

Calculate a two-sided permutation-based p-value. In other words, count the
number of test statistics generated from the permutations with a mean difference
that is greater in magnitude that the observed difference, and divide by the
number of permutations.

```{r p-value}
p_perm <- (sum(diff_permutations <= -abs(diff_observed)) +
           sum(diff_permutations >= abs(diff_observed))) / length(diff_permutations)
p_perm
```


The p-value is `r p_perm`.
